Using SIMD and SIMT vectorization to evaluate sparse chemical kinetic Jacobian matrices and thermochemical source terms

### 2018-06-04 paper.tex
a few more edits in page 7

A number of recent works have investigated \revise{using} high-performance SIMT devices \revise{like GPUs} to accelerate reactive-flow and chemical kinetics simulations.
Spafford et al.~\cite{Spafford:2010aa} coupled GPU-based chemical source-term evaluation with an explicit direct numerical simulation code, achieving an order of magnitude speedup \delete{on a Tesla C1060} compared to a \add{CPU-based} serial implementation \delete{on a AMD-Opteron processor}.
Shi et al.~\cite{Shi:2011aa} combined GPU-based chemical kinetic source-term evaluation and Jacobian factorization with two implicit CPU solvers, achieving an order-of-magnitude speedup for homogeneous reactor simulations of large chemical models\delete{on a Tesla 2050 GPU} over a serial\delete{Intel i7 930} CPU implementation.
Niemeyer et al.~\cite{Niemeyer:2011aa} implemented an explicit solver for non-stiff chemistry on a Tesla C2075 GPU, achieving a speedup of nearly two orders of magnitude over a sequential \add{CPU} code\delete{on an Intel Xeon X5650 CPU}.
Shi et al.~\cite{Shi:2012aa} proposed a strategy for chemical-kinetic integration in three-dimensional reactive-flow simulations, where a traditional implicit integrator handled the stiffest computational cells on a CPU and a stabilized-explicit solver \delete{on a Tesla C2050 GPU} solved the less-stiff cells on a GPU; this hybrid solution technique \revise{performs} \SIrange{11}{46}{$\times$} faster than the implicit CPU solver alone for simulation of a premixed diesel engine.
Le et al.~\cite{Le2013596} found a \SIrange{30}{50}{$\times$} speedup for a GPU-based shock-capturing reactive-flow code\delete{ on a Tesla C2070} as compared with a sequential \add{CPU} version of the same\delete{on an Intel Xeon X5650 CPU}.
Stone and Davis~\cite{Stone:2013aa} investigated a GPU-based version of a common implicit integrator (VODE~\cite{Brown:1989vl}), finding an order-of-magnitude speedup over a serial CPU implementation.
Niemeyer and Sung~\cite{Niemeyer:2014aa} developed a GPU-based stabilized explicit integrator for use with moderately-stiff chemical kinetics, achieving an order of magnitude speedup\delete{on a Tesla C2075 GPU} over a multithreaded VODE solver on a six-core\delete{Intel X5650} CPU.
Sewerin and Rigopoulos~\cite{Sewerin20151375} studied a fifth-order implicit Runge--Kutta solver on both consumer-grade and high-end GPUs\slash CPUs; the high-end GPU solver was at best \SI{1.8}{$\times$} slower than the high-end CPU version running on \num{16} cores.
Yonkee and Sutherland~\cite{Yonkee2016} implemented GPU-accelerated evaluations of thermodynamic parameters, multicomponent transport properties, and species production rates for solving partial differential equations (PDEs) \delete{on an Intel Xeon E5-2680} \add{both the CPU} and \delete{a Tesla K20} GPU.
In evaluating the thermochemical properties, speedups between \SIrange{8}{13}{$\times$} were found (over a serial version of the same code) while running on 16 CPU cores, and \SIrange{20}{40}{$\times$} on the GPU; speedups of \textasciitilde\SI{9}{$\times$} and \textasciitilde\SI{25}{$\times$} were found in solution of a partially premixed methanol flame PDE on 16 CPU cores and the GPU, respectively.\todo{revise this sentence, fairly awkward}{}
Curtis et al.~\cite{CurtisGPU:2017} implemented a fifth-order implicit Runge--Kutta method~\cite{wanner1991solving}, as well as two fourth-order exponential integration techniques~\cite{Hochbruck:1998,Hockbruck:2009} paired with an analytical Jacobian code~\cite{Niemeyer:2016aa}\delete{on a Tesla C2075} \add{on the GPU and CPU}.
The \add{GPU-based} implicit Runge--Kutta method was roughly equivalent to a standard implicit integrator~\cite{Hindmarsh:2005} running on \numrange{12}{38}\delete{Intel Xeon E5-4640 v2} CPU cores for two relatively small chemical models \delete{with an integration time step of $10^{-6}$ s}.

>>>>

A number of recent works have investigated \revise{using} high-performance SIMT devices \revise{like GPUs} to accelerate reactive-flow and chemical kinetics simulations.
Spafford et al.~\cite{Spafford:2010aa} coupled GPU-based chemical source-term evaluation with an explicit direct numerical simulation code, achieving an order of magnitude speedup \delete{on a Tesla C1060} compared to a \add{CPU-based} serial implementation \delete{on a AMD-Opteron processor}.
Shi et al.~\cite{Shi:2011aa} combined GPU-based chemical kinetic source-term evaluation and Jacobian factorization with two implicit CPU solvers, achieving an order-of-magnitude speedup for homogeneous reactor simulations of large chemical models\delete{on a Tesla 2050 GPU} over a serial\delete{Intel i7 930} CPU implementation.
Niemeyer et al.~\cite{Niemeyer:2011aa} implemented an explicit solver for non-stiff chemistry on a Tesla C2075 GPU, achieving a speedup of nearly two orders of magnitude over a sequential \add{CPU} code\delete{on an Intel Xeon X5650 CPU}.
Shi et al.~\cite{Shi:2012aa} proposed a strategy for chemical-kinetic integration in three-dimensional reactive-flow simulations, where a traditional implicit integrator handled the stiffest computational cells on a CPU and a stabilized-explicit solver \delete{on a Tesla C2050 GPU} solved the less-stiff cells on a GPU; this hybrid solution technique \revise{performs} \SIrange{11}{46}{$\times$} faster than the implicit CPU solver alone for simulation of a premixed diesel engine.
Le et al.~\cite{Le2013596} found a \SIrange{30}{50}{$\times$} speedup for a GPU-based shock-capturing reactive-flow code\delete{ on a Tesla C2070} as compared with a sequential \add{CPU} version of the same\delete{on an Intel Xeon X5650 CPU}.
Stone and Davis~\cite{Stone:2013aa} investigated a GPU-based version of a common implicit integrator (VODE~\cite{Brown:1989vl}), finding an order-of-magnitude speedup over a serial CPU implementation.
Niemeyer and Sung~\cite{Niemeyer:2014aa} developed a GPU-based stabilized explicit integrator for use with moderately-stiff chemical kinetics, achieving an order of magnitude speedup\delete{on a Tesla C2075 GPU} over a multithreaded VODE solver on a six-core\delete{Intel X5650} CPU.
Sewerin and Rigopoulos~\cite{Sewerin20151375} studied a fifth-order implicit Runge--Kutta solver on both consumer-grade and high-end GPUs\slash CPUs; the high-end GPU solver was at best \SI{1.8}{$\times$} slower than the high-end CPU version running on \num{16} cores.
Yonkee and Sutherland~\cite{Yonkee2016} implemented GPU-accelerated evaluations of thermodynamic parameters, multicomponent transport properties, and species production rates for solving partial differential equations (PDEs) \delete{on an Intel Xeon E5-2680} \add{both the CPU} and \delete{a Tesla K20} GPU.
In evaluating the thermochemical properties, speedups between \SIrange{8}{13}{$\times$} were found (over a serial version of the same code) while running on 16 CPU cores, and \SIrange{20}{40}{$\times$} on the GPU; they showed speedups of \textasciitilde\SI{9}{$\times$} and \textasciitilde\SI{25}{$\times$} when simulating a partially premixed methanol flame PDE on 16 CPU cores and the GPU, respectively.\todo{revise this sentence, fairly awkward}{}
Curtis et al.~\cite{CurtisGPU:2017} implemented a fifth-order implicit Runge--Kutta method~\cite{wanner1991solving}, as well as two fourth-order exponential integration techniques~\cite{Hochbruck:1998,Hockbruck:2009} paired with an analytical Jacobian code~\cite{Niemeyer:2016aa} \delete{on a Tesla C2075} \add{on the GPU and CPU}.
The \add{GPU-based} implicit Runge--Kutta method performs equivalently to a standard implicit integrator~\cite{Hindmarsh:2005} running on \numrange{12}{38} \delete{Intel Xeon E5-4640 v2} CPU cores for two relatively small chemical models \delete{with an integration time step of $10^{-6}$ s}.
